Dynamics of Laterally Propagating Flames in X-ray Bursts. II. Realistic Burning \& Rotation

### 2020-12-17 xrb2/paper.tex
made phase plot bullets into paragraphs

Figures~\ref{fig:time_series_500} and \ref{fig:time_series_1000} show time series of the mean molecular weight ($\bar{A}$) for the {\tt F500} and {\tt F1000} runs.  Compared to those in \cite{flame_wave1}, ashes behind the flame do not reach as high atomic weights. This is not surprising, since those early runs boosted the reaction rates.  Comparing these two new runs, the burning is much more evolved for the higher rotation rate.  We believe that this is because the increased rotation confines the initial perturbation and subsequent expansion from the burning better, allowing the reactions to progress further. In contrast to {\tt F500} and {\tt F1000}, the lowest rotation run --- {\tt F250} --- failed to ignite. The lack of ignition for {\tt F250} also aligns with the reasoning given above, with the lower rotation in this case potentially leading to insufficient confinement of the burning. In this scenario, another source of confinement (e.g.\ magnetic fields) would need to take over at lower rotation rates to allow a burst to occur, at least for our current initial model.

Looking at the phase plots in Figures~\ref{fig:urho}-\ref{fig:abar}, we see:
\begin{itemize}
    \item From the $u$-$\rho$ and $u$-$v$ plots, we see that the burning occurs in a region of high density where the fluid is moving in a circular motion. From the previous plots (see: streamlines?), this most likely corresponds to the leading edge of the flame. This feature is not (yet?) developed in the 250~Hz flame. (will this ever develop, or will it just fizzle out completely?)
    \item From the $\bar{A}$-$T$ plots in Figure \ref{fig:abar}, we can see that as the rotation rate increases, so does the peak temperature. We believe that the tracks in the plot correspond to reaction pathways in the reaction network. In the 1000~Hz plot, these tracks are somewhat disrupted, possibly due to the more vigorous mixing of the vortex at the flame front. (If this continues or if the rotation was cranked even higher, would the reactions get so disrupted that the burning is snuffed out, similar to what happens in combustion when the turbulent vortices become so small they penetrate the reaction zone and completely disrupt it?)
\end{itemize}

\begin{figure}[t]
    \centering
    \plotone{urho_plot_0_1}
    \caption{\label{fig:urho}Phase plot of the horizontal $x$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 250, 500 and 1000~Hz runs.}
\end{figure}

>>>>

The {\tt F500} and {\tt F1000} runs both qualitatively resemble the flame structure in \citet{flame_wave1} --- a laterally propagating flame that is lifted off of the bottom of the fuel layer --- but they differ in their burning structures. Figures~\ref{fig:time_series_500} and \ref{fig:time_series_1000} show time series of the mean molecular weight ($\bar{A}$) for the {\tt F500} and {\tt F1000} runs.  Compared to those in \cite{flame_wave1}, ashes behind the flame do not reach as high atomic weights. This is not surprising, since those early runs boosted the reaction rates.  Comparing these two new runs, the burning is much more evolved for the higher rotation rate, and the ash is actually able to move ahead of the flame front (visible in the Figure  \ref{fig:time_series_1000} 100 ms snapshot). We believe that this is because the increased rotation confines the initial perturbation and subsequent expansion from the burning better, allowing the reactions to progress further. In contrast to {\tt F500} and {\tt F1000}, the lowest rotation run --- {\tt F250} --- failed to ignite. The lack of ignition for {\tt F250} also aligns with the reasoning given above, with the lower rotation in this case potentially leading to insufficient confinement of the burning. In this scenario, another source of confinement (e.g.\ magnetic fields) would need to take over at lower rotation rates to allow a burst to occur, at least for our current initial model.

Burning in the {\tt F500} and {\tt F1000} runs is concentrated in a dense region with circular motion. Figure \ref{fig:urho} compares the horizontal $x$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for {\tt F250}, {\tt F500}, and {\tt F1000}. Most of the burning for each of the simulations occurs in a high density region $\rho > 3 \times 10^5 {\rm g/cm^3}$. The fluid in this dense, high energy generation region undergoes vortical motion, shown in the Figure \ref{fig:uv} phase plots comparing $u$, $v$ and $\dot{e}_{\rm nuc}$. From the previous plots (see: streamlines?) \MarginPar{NF: which plots?}, this most likely corresponds to the leading edge of the flame where fresh fuel is being entrained. This feature is not developed in the 250~Hz flame; it could potentially develop at later times, or the flame could just fizzle out.

The mean molecular weight $\bar{A}$ within each of our simulations seems to grow along defined tracks confined to certain temperatures $T$, as shown in the Figure \ref{fig:abar} phase plots. We believe that the tracks in the plot correspond to reaction pathways in the reaction network. Comparing Figure \ref{fig:abar} to \citet{flame_wave1}, these tracks are much more neat and clearly defined. The ''messiness" of the tracks may be dependent on how mixed the flame interior is --- since these new simulations are un-boosted, they may be inherently less mixed than those in \citet{flame_wave1}. {\tt F1000} aligns with this interpretation: its $\bar{A}$ tracks are somewhat disrupted compared to the slower rotation runs, possibly due to the more vigorous mixing of the vortex at the flame front. Comparing the different runs, we also see that as the rotation rate increases, so does the peak temperature. This makes sense if higher rotation leads to a more confined, intense vortex near the flame front. It also agrees with our earlier interpretation of enhanced burning seen in Figure \ref{fig:time_series_1000} for {\tt F1000}. The disruption of reaction pathways with higher rotation rates and mixed flames may have interesting consequences; if the reactions are disrupted enough, it is possible that the burning could be snuffed out. This can occur in combustion scenarios when the turbulent vortices become so small they penetrate the reaction zone and completely disrupt the burning.

\begin{figure}[t]
    \centering
    \plotone{urho_plot_0_1}{}
    \caption{\label{fig:urho}Phase plot of the horizontal $x$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 250, 500 and 1000~Hz runs.}
\end{figure}
